---
title: "README"
author: "Spenser Leise"
date: "2024-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Examining 'Hot Ones' Data
Luke Juffer, Ryan Nagao, Mason Sackpraseuth, and Spenser Leise

## Introduction:

  Hot Ones is a popular internet show about guests from various industries eating spicy wings that get increasingly hot wing after wing. The guest is also interviewed as the show goes on. We have 3 sets of data that have been scraped off of the Hot Ones Wikipedia. Hot Ones has become a cultural landmark in recent years, and we believe that by analyzing the data available we can draw interesting conclusions regarding celebrity status and viewer's interests.
  
## Mission Statement:
  
  To find various correlations through the 3 data sets that reveal key insights to the show, with hopes to find answers for questions of various magnitudes. We want to expand on the data as well as draw correlations from what is present. 
  
## Importing the data sets

  There are 3 data sets that make up the hot ones data set. all of them being important to our mission. One is episodes, the people featured and data if they finished their wings or not. Then Sauces, where we can see how hot each sauce is and the name. Then Seasons which has other notes. 

```{r}
episodes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-08-08/episodes.csv')
sauces <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-08-08/sauces.csv')
seasons <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-08-08/seasons.csv')
```
```{r}
head(seasons)
```
```{r}
head(sauces)
```

```{r}
head(episodes)
```
## Potential questions we could ask
How many people have failed the show?
What season has the hottest sauce?
What is the hottest sauce featured on the show?
What industries have the most people on hot ones? Are there any correlations to their performance?
What correlations can be made for each season, and what changes were made?
What observable changes are there on the types of guests that appear over time?

## Cleaning The data/preparing data
As our data sets are scraped off the web, it would be assumed that there would be viggorous cleaning needed for the set itself. However, Tidy Tuesday helps us with a cleaning script that had already been ran on the data. We have included the cleaning script here (it will fail as it has already been ran with the package)
```{r}
### Suggested TidyTuesday dataset: 
### 300 Hot Ones episodes (guests and sauces), seasons 1-21 (current)
### Original script by Carl Börstell (@borstell)

#https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-08-08/readme.md
```

There are a couple of columns that are not needed for our exploration. The notes column in the season's table is not important to our exploration. We removed all this with a cleaning script of our own. Shown Below.
```{r}
library(tidyverse)
library(rvest)
library(here)
library(data.table)
seasons <- seasons %>%
  select(-note)

```
## Initial Data Observation
```{r}
summary(episodes)

summary(sauces)

summary(seasons)
```
**The Longest data set is the episodes one, with 300 individual entries. We can also see that the show has been active for around 8 years and counting. The finished column also tells us how many people have failed the show, with 15 failing to eat all the wings. This is much lower than I personally expected. We can also observe the hottest sauce on the show. The highest scoville sauce was 2000000 units which is  880 times hotter than your garden variety jalapeño.**

# Variable Names
## episodes
season: what season the episode premiered during
episode_overall: what number the episode is, regardless of season
episode_season: what number the episode is within the season
title: the title of the youtube video
original_release: release date of the youtube video
guest: the name of the celebrity guest
guest_appearance_number: the number of times the guest had been on the show
finished: false if they did not finish their wings, true if they did
occupation: what field the guest works in

## sauces
season: what season the sauce appeared during
sauce_number: what order the sauces were presented in during the episode
sauce_name: the name of the sauce
scoville: a measurement of the spice level of the sauce

## seasons
season: what season the sauce appeared during
episodes: number of episodes in the season
original_release: release date of the first episode in the season
last_release: release date of the last episode in the season

# Data Exploration

### Spenser's Exploration
I was interested in the failure rate of each season and how where a person comes from can affect their spice tolerance. I had to add a new column of all the industries that each person is from and then take averages of the failure rate.
```{r}
#new occupation column 
occupation <- c(
    "Music", "Sports", "Music", "Music", "Music", "Music", "Sports", "Music", "Music", "Music",
    "Acting", "Acting", "Comedy", "Music", "Acting", "Comedy", "Music", "Comedy", "Comedy", "Comedy",
    "Comedy", "Comedy", "Sports", "Sports", "Comedy", "Comedy", "Comedy", "Comedy", "Music", "Comedy",
    "Music", "Comedy", "Acting", "Comedy", "Sports", "Comedy", "Comedy", "Comedy", "Acting", "Acting",
    "Sports", "Music", "Music", "Acting", "Comedy", "Acting", "Comedy", "Comedy", "Comedy", "Acting",
    "Music", "Comedy", "Comedy", "Acting", "Comedy", "Comedy", "Acting", "Comedy", "Entertainment", "Acting",
    "Acting", "Acting", "Sports", "Sports", "Entertainment", "Acting", "Comedy", "Comedy", "Entertainment", "Acting",
    "Comedy", "Comedy", "Music", "Sports", "Acting", "Acting", "Sports", "Music", "Acting", "Music",
    "Music", "Acting", "Comedy", "Comedy", "Comedy", "Music", "Music", "Sports", "Comedy", "Acting",
    "Acting", "Acting", "Music", "Entertainment", "Entertainment", "Acting", "Sports", "Music", "Comedy", "Acting",
    "Acting", "Acting", "Comedy", "Acting", "Music", "Comedy", "Music", "Comedy", "Acting", "Music",
    "Acting", "Comedy", "Entertainment", "Comedy", "Acting", "Sports", "Music", "Comedy",
    "Comedy", "Acting", "Music", "Comedy", "Acting", "Comedy", "Acting", "Comedy", "Music", "Music",
    "Acting", "Acting", "Acting", "Acting", "Acting", "Acting", "Music", "Entertainment",
    "Acting", "Acting", "Acting", "Music", "Acting", "Acting", "Music", "Music", "Acting", "Acting",
    "Comedy", "Acting", "Music", "Acting", "Acting", "Acting", "Music", "Acting", "Comedy", "Acting",
    "Music", "Acting", "Acting", "Acting", "Acting", "Music", "Acting", "Acting", "Acting", "Comedy",
    "Acting", "Sports", "Acting", "Acting", "Acting", "Acting", "Acting", "Music", "Acting",
    "Entertainment", "Acting", "Comedy", "Sports", "Acting", "Acting", "Acting", "Acting",
    "Music", "Acting", "Acting", "Acting", "Acting", "Acting", "Acting", "Music", "Acting",
    "Acting", "Acting", "Acting", "Acting", "Acting", "Music", "Acting", "Acting", "Acting",
    "Acting", "Acting", "Acting", "Music", "Acting", "Acting", "Comedy", "Music", "Sports", "Acting",
    "Acting", "Sports", "Acting", "Sports", "Acting", "Acting", "Acting", "Acting", "Acting", "Comedy",
    "Comedy", "Acting", "Acting", "Comedy", "Entertainment", "Music", "Acting", "Comedy",
    "Acting", "Music", "Acting", "Acting", "Acting", "Acting", "Acting", "Entertainment", "Music", "Acting",
    "Acting", "Acting", "Comedy", "Music", "Acting", "Acting", "Acting", "Acting", "Acting", "Music",
    "Comedy", "Acting", "Acting", "Acting", "Acting", "Comedy", "Acting", "Acting", "Acting", "Acting",
    "Acting", "Acting", "Music", "Acting", "Acting", "Acting", "Acting", "Acting", "Acting", "Acting",
    "Music", "Acting", "Acting", "Acting", "Acting", "Music", "Acting", "Comedy", "Music", "Sports",
    "Acting", "Sports", "Acting", "Acting", "Acting", "Acting", "Acting", "Acting", "Acting", "Acting",
    "Music", "Acting", "Entertainment", "Comedy", "Acting", "Sports", "Acting", "Acting", "Music",
    "Acting", "Sports"
)

episodes <- episodes %>%
  mutate(occupation = occupation)

# Display the updated 'Episodes' data frame
head(episodes)
```

You can also embed plots, for example:
**Spenser Leise Question 1: What Occupation had the highest failure rate**
```{r}
library(dplyr)

# Calculate failure rate by occupation
failure_rates <- episodes %>%
  group_by(occupation) %>%
  summarise(failure_rate = mean(!finished))

library(ggplot2)

failure_rates$color <- cut(failure_rates$failure_rate, breaks = c(0.03, 0.04, 0.05, 0.06, 1),
                           labels = c("#5EFF33","#DAF7A6", "#FFC300", "#FF5733"))

# Create a bar chart
ggplot(failure_rates, aes(x = reorder(occupation, -failure_rate), y = failure_rate, fill = color)) +
  geom_bar(stat = "identity", width = 0.7) +
  labs(
    title = "Hot Ones Failure Rate by Occupation",
    x = "Occupation",
    y = "Failure Rate",
    caption = "Over the course of first 300 episodes"
  ) +
  scale_fill_identity() +  # Use custom fill colors
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**what season was the hottest?**
```{r}
failure_rates <- episodes %>%
  group_by(season) %>%
  summarise(failure_rate = mean(!finished)) %>%
  arrange(season)  # Sort by season (optional)

# Create a bar chart
ggplot(failure_rates, aes(x = factor(season), y = failure_rate, fill = factor(season))) +
  geom_bar(stat = "identity", width = 0.7) +
  labs(
    title = "Failure Rate by Season",
    x = "Season",
    y = "Failure Rate",
    caption = "Based on sample data"
  ) +
  scale_fill_discrete(name = "Season") +  # Use discrete fill scale for season
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}

scoville_avg <- sauces%>%
  group_by(season) %>%
  summarise(avg_scoville = mean(scoville))

# Create a line plot of average Scoville rating by season
ggplot(scoville_avg, aes(x = season, y = avg_scoville)) +
  geom_line(color = "orange", size = 1.5) +
  geom_point(color = "red", size = 3) +
  labs(
    title = "Average Scoville Rating by Season",
    x = "Season",
    y = "Average Scoville Rating",
    caption = "Based on sample data"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = scoville_avg$season, labels = scoville_avg$season) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Luke Juffer's exploration
I was interested in what occupations show up most frequently on hot ones how the length of seasons compare.

**what occupations show up most frequently?**
```{r}
ggplot(data = episodes, aes(x = occupation)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Comparison of Occupations",
       x = "Occupation",
       y = "Count")


ggplot(data = episodes, aes(x = occupation)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Comparison of Occupations",
       x = "Occupation",
       y = "Count") +
  facet_wrap(~ season) +
  theme(axis.text.x = element_blank(),  
        strip.text.x = element_blank()) 

```
<br>**It seems the amount of actors appearing on the show has increased relative to other occupations as the seasons progress. Comedy and entertainment made up a significant portion of guests in the early seasons, but seem to have dropped off significantly after season 6 or so. Sports seems to be the most consistently lowest appearing occupation.**

**How do the lengths of seasons compare?**
```{r}
seasons$length <- seasons$last_release - seasons$original_release
head(seasons)

ggplot(seasons, aes(x = season, y = length, group = 1)) +
  geom_line(color = "blue") +  
  geom_point(color = "blue") +  
  labs(x = "Season Number", y = "Length") +  
  theme_minimal() +  
  ggtitle("Length Over Time")  
```
<br>**The first and second season were much longer than the others, with a sharp decline occurding after season 2 into season 3. The longest season was season 2, being around 400 days long. Length of season stabilizes after season 5 at around 90 days.**


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### Ryan's Exploration
Viewers of the show will recognize the notorious sauce "The Last Dab", but we wanted to verify that it was the hottest sauce of all.
**what is the hottest sauce?**
```{r }
hottest_sauces <- sauces |>
  filter(scoville == max(scoville)) |>
  select(sauce_name)

unique(hottest_sauces$sauce_name)
print(max(sauces$scoville))
```
**The Last Dab and all its variations are, indeed, the hottest sauce. They all have a heat of 2 million scoville units.**

```{r }
hottest_seasons <- sauces |>
  filter(scoville == max(scoville)) |>
  select(season)
unique(hottest_seasons$season)
```
**From Season 4 onwards, the Last Dab (or one of its variations) appears.**

**How many people have failed Hot Ones?**
```{r }
episodes |>
  filter(finished == FALSE) |>
  select(season, guest)
```
**Fifteen people have failed hot ones: Tony Yayo, DJ Khaled, Mike Epps, Jim Gaffigan, Rob Corddry, Ricky Gervais, Mario Batali, Taraji P. Henosn, Lil Yachty, E-40, Shaq, Chance the Rapper, Eric André, Quavo, and Pusha T.**

# Conclusions
In conclusion, we have drawn out several interesting insights from our datasets regarding Hot Ones and it's guests. For example, we found that the failure rate was the highest by far in season 1 at 25%, dropping off sharply after and hovering at around 8%. This could suggest that cultural or viewership changes took place, and that guests on the show don't wish to appear uncommitted or weak by not finishing their wings. This is reinforced by the fact that we see the average scoville rating by season increase to 300,000 by season 5. We would expect this to increase the failure rate, but the fact that failure rate remains low suggests that external factors influence the failure rate. We also observe a sharp decline in the length of seasons between 1-5. These observations suggest that the show did not reach a steady formula until around the 5th season.

Further data collection could allow us to draw out additional interesting conclusions. For example, data on views per video would let us look at the effect of occupation, release date, and maybe even title on viewership. We could also collect data for more recent episodes, as episodes are constatly being released.